---
title: |
  | Summarizing Distributions of
  | Latent Structures
author: |
  | David B. Dahl: Brigham Young University
  | Peter Müller: University of Texas at Austin
  |
  | Bayesian Nonparametric Inference: Dependence Structures & Applications
  | Oaxaca, Mexico
  | December 6, 2017
#date: "December 6, 2017"

fontsize: 11pt
output:
  beamer_presentation:
    theme: "metropolis"
    #colortheme: "dolphin"
    fonttheme: "structurebold"
    includes:
      in_header: latex-topmatter.tex
---

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
knitr::opts_chunk$set(cache=TRUE)
options(width = 50)

library(sdols)
iris.clusterings <- unname(iris.clusterings)
tmp <- iris.clusterings[1,]
iris.clusterings[1,] <- iris.clusterings[13,]
iris.clusterings[13,] <- tmp
```

## Typical Bayesian data analysis workflow...

- Perform exploratory data analysis to understand the data.
- Define sampling model and prior distribution.
- Program & debug the model implementation.
- Obtain posterior draws by letting the MCMC run a long time.
- Calculate Bayes estimates by summarize posterior draws.
    - Real- and vector-valued parameters... compute mean
    - **Latent structures (e.g., partitions, feature allocations, networks)... ???**
- **Report results**
    
## Example: Dirichlet Mixture Model with Gaussian Likelihood

Suppose MCMC output contains 1,000 posterior samples for the partition of 150 observations (encoded with cluster labels), e.g.:
```{r, echo=TRUE}
iris.clusterings[1,]
```
What is the Bayes estimate of the partition from these 1,000 samples?

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

This is very important.
```{r conf, echo=TRUE}
library(sdols)
ppm <- expectedPairwiseAllocationMatrix(iris.clusterings)
est <- salso(ppm)
conf <- confidence(est,ppm)
#plot(conf)
plot(conf,data=iris)
```

## Another slide

It is very important $\sum_i^n x^2$.

$$\sum_{i=1}^n (x_i - \bar{x})^2$$

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

